# Why?

Feature checklist

* **Track** project history
* **Undo** infinitely
* **Branch** what-if scenarios (without breaking primary analysis)
* **Sync** local workstation with other computers
* **Share** with collaborators
  * Most important collaborators: past-self and future-self
* **Package** project for a code repository


# Typical workflow

1. Start on local workstation
  1. Initialize and setup a project
  1. Rapidly develop on simulated data or a small subset of data
  1. Test code (as opposed to test analysis)
  1. Test-revise until ready to...
1. Scale up on Exacloud/Norm
  1. Run analysis on full data
  1. Revise analysis
  1. Generate and collect analytic products for...
1. Synthesis on local workstation
  1. Refine analysis into a final product
  1. Handoff to stakeholder
  1. Package entire project into a repository for archive

```{r diagramNonGitWorkflow, echo = FALSE, fig.height = 2.5}
G <-
  create_graph(graph_attrs = c("rankdir = LR"),
               node_attrs = c("fontname = \"Helvetica\"")) %>%
  add_node(label = "Local") %>%
  add_node(label = "Exacloud/Norm") %>%
  add_edge(1, 2) %>%
  add_edge(2, 1)
render_graph(G)
```

I could simply copy-paste or FTP files to make this work.
However, that only checks the **Sync** feature, and inelegantly at that.

If the data is small and I don't need to scale up, then I can skip the **Sync** feature.
But I still need to check the other features.


# Install git

1. Install [Git](https://git-scm.com/) on local workstation
   * Includes a GUI tool
1. *Optional* Install a 3rd-party [GUI client](https://git-scm.com/downloads/guis)
1. Exacloud/Norm already has git installed


# Initialize a local project repo

```{r, echo = FALSE}
G <-
  create_graph(graph_attrs = c("rankdir = LR"),
               node_attrs = c("fontname = \"Helvetica\""),
               edge_attrs = c("fontname = \"Helvetica\""))
G <-
  G %>%
  add_node(label = "Local") %>%
  add_node(label = "Bare repo\na.k.a. origin") %>%
  set_node_attrs(nodes = 2, "color", "grey80") %>%
  set_node_attrs(nodes = 2, "fontcolor", "grey80") %>%
  add_node(label = "Exacloud/Norm") %>%
  add_edge(1, 2) %>% set_edge_attrs(from = 1, to = 2, "label", "push") %>%
  add_edge(2, 1) %>% set_edge_attrs(from = 2, to = 1, "label", "pull") %>%
  add_edge(2, 3) %>% set_edge_attrs(from = 2, to = 3, "label", "pull") %>%
  add_edge(3, 2) %>% set_edge_attrs(from = 3, to = 2, "label", "push")
```

From your local Git Bash command line:

```
$ git init
```

```{r, echo = FALSE, fig.height = 2.5}
colAction <- "green"
colHide <- rgb(0, 0, 0, 0)
G <-
  G %>%
  set_edge_attrs(from = c(1, 2, 2, 3), to = c(2, 1, 3, 2), "color", colHide) %>%
  set_edge_attrs(from = c(1, 2, 2, 3), to = c(2, 1, 3, 2), "fontcolor", "white") %>%
  set_node_attrs(nodes = c(2, 3), "color", colHide) %>%
  set_node_attrs(nodes = c(2, 3), "fontcolor", "white") %>%
  set_node_attrs(nodes = 1, "color", colAction) %>%
  set_node_attrs(nodes = 1, "fontcolor", colAction)
render_graph(G)
G <-
  G %>%
  set_node_attrs(nodes = 1, "color", "black") %>%
  set_node_attrs(nodes = 1, "fontcolor", "black")
```


# Bare repositories

To make the connection to **Sync** to/from, we need to create something called a **bare repository**.

Bare repositories can exist anywhere.
E.g., on a local drive, on the Exacloud/Norm Linux filesystem, [GitHub](https://github.com).
The only constraint is that your local workstation and your remote computer need to be able to connect to it.

Bare repo location | Pros | Cons
:---|:---|:---
Local | Private | Exacloud/Norm won't be able to connect to it
[Box.com](https://ohsu.account.box.com) | Shareable | Exacloud/Norm won't be able to connect to it
Home directory on Exacloud/Norm `~` | Private, behind OHSU firewall |
Group directory on Exacloud/Norm `/home/groups/biostats` | Shareable, behind OHSU firewall |
[GitHub](https://github.com) | Shareable to anyone | In the cloud

Techincally speaking, a bare repository is a repo that doesn't contain any working files.
Practically speaking, once a bare repo is set up, it's invisible to the user.


# Initialize a bare repo 

From the Linux command line on Exacloud/Norm:

1. Create a directory with the `.git` extension

   ```
   $ mkdir <repo-name>.git
   ```

1. Navigate into the directory

   ```
   $ cd <repo-name>.git
   ```

1. Initialize a bare repository

   ```
   $ git init --bare
   ```

```{r, echo = FALSE, fig.height = 2.5}
G <-
  G %>%
  set_node_attrs(nodes = 2, "color", colAction) %>%
  set_node_attrs(nodes = 2, "fontcolor", colAction)
render_graph(G)
G <-
  G %>%
  set_node_attrs(nodes = 2, "color", "grey80") %>%
  set_node_attrs(nodes = 2, "fontcolor", "grey80")
```


# Connect local repo to bare

From your local Git Bash command line:

```
$ git remote add origin <username>@exacloud.ohsu.edu:<repo-name>.git
```

```{r, echo = FALSE, fig.height = 2.5}
G <-
  G %>%
  set_edge_attrs(from = 1, to = 2, "color", colAction) %>%
  set_edge_attrs(from = 2, to = 1, "color", colAction)
render_graph(G)
G <-
  G %>%
  set_edge_attrs(from = 1, to = 2, "color", "grey80") %>%
  set_edge_attrs(from = 2, to = 1, "color", "grey80")
```


# Push your local repo

From your local Git Bash command line:

```
$ git push origin master
```

```{r, echo = FALSE, fig.height = 2.5}
G <-
  G %>%
  set_edge_attrs(from = 1, to = 2, "color", colAction) %>%
  set_edge_attrs(from = 1, to = 2, "penwidth", 2) %>%
  set_edge_attrs(from = 1, to = 2, "fontcolor", colAction)
render_graph(G)
G <-
  G %>%
  set_edge_attrs(from = 1, to = 2, "color", "black") %>%
  set_edge_attrs(from = 1, to = 2, "penwidth", 1) %>%
  set_edge_attrs(from = 1, to = 2, "fontcolor", "black")
```


# Initialize a project repo on Exacloud/Norm filesystem

From the Linux command line on Exacloud/Norm:

```
$ git init
```

```{r, echo = FALSE, fig.height = 2.5}
G <-
  G %>%
  set_node_attrs(nodes = 3, "color", colAction) %>%
  set_node_attrs(nodes = 3, "fontcolor", colAction)
render_graph(G)
G <-
  G %>%
  set_node_attrs(nodes = 3, "color", "black") %>%
  set_node_attrs(nodes = 3, "fontcolor", "black")
```


# Connect Exacloud/Norm repo to bare

From the Linux command line on Exacloud/Norm:

```
$ git remote add origin <repo-name>.git
```

```{r, echo = FALSE, fig.height = 2.5}
G <-
  G %>%
  set_edge_attrs(from = 2, to = 3, "color", colAction) %>%
  set_edge_attrs(from = 3, to = 2, "color", colAction)
render_graph(G)
G <-
  G %>%
  set_edge_attrs(from = 2, to = 3, "color", "grey80") %>%
  set_edge_attrs(from = 3, to = 2, "color", "grey80")
```


# Pull to your Exacloud/Norm repo

From the Linux command line on Exacloud/Norm:

```
$ git pull origin master
```

```{r, echo = FALSE, fig.height = 2.5}
G <-
  G %>%
  set_edge_attrs(from = 2, to = 3, "color", colAction) %>%
  set_edge_attrs(from = 2, to = 3, "penwidth", 2) %>%
  set_edge_attrs(from = 2, to = 3, "fontcolor", colAction)
render_graph(G)
G <-
  G %>%
  set_edge_attrs(from = 2, to = 3, "color", "black") %>%
  set_edge_attrs(from = 2, to = 3, "penwidth", 1) %>%
  set_edge_attrs(from = 2, to = 3, "fontcolor", "black")
```


# Create, stage, commit in Exacloud/Norm repo

Create a `Hello world` file.
Stage `test.txt` using `git add`.
Commit changes with a meaningly message using `git commit`.
Add another line to `test.txt`.
Stage and commit changes.
Check `git status` liberally.

```
$ git status
$ echo "Hello world. This is a test file." > test.txt
$ git status
$ git add test.txt
$ git status
$ git commit -m "Create hello world file"
$ git status
$ echo "Created by me." >> test.txt
$ git status
$ git add test.txt
$ git status
$ git commit -m "Add attribution"
$ git status
```


# Push your Exacloud/Norm repo

From the Linux command line on Exacloud/Norm:

```
$ git push origin master
```

```{r, echo = FALSE, fig.height = 2.5}
G <-
  G %>%
  set_edge_attrs(from = 3, to = 2, "color", colAction) %>%
  set_edge_attrs(from = 3, to = 2, "penwidth", 2) %>%
  set_edge_attrs(from = 3, to = 2, "fontcolor", colAction)
render_graph(G)
G <-
  G %>%
  set_edge_attrs(from = 3, to = 2, "color", "black") %>%
  set_edge_attrs(from = 3, to = 2, "penwidth", 1) %>%
  set_edge_attrs(from = 3, to = 2, "fontcolor", "black")
```


# Pull to your local repo

From your local Git Bash command line:

```
$ git pull origin master
```

```{r, echo = FALSE, fig.height = 2.5}
G <-
  G %>%
  set_edge_attrs(from = 2, to = 1, "color", colAction) %>%
  set_edge_attrs(from = 2, to = 1, "penwidth", 2) %>%
  set_edge_attrs(from = 2, to = 1, "fontcolor", colAction)
render_graph(G)
G <-
  G %>%
  set_edge_attrs(from = 2, to = 1, "color", "black") %>%
  set_edge_attrs(from = 2, to = 1, "penwidth", 1) %>%
  set_edge_attrs(from = 2, to = 1, "fontcolor", "black")
```


# Complete workflow

```{r diagramGitWorkflow, echo = FALSE, fig.height = 2.5}
render_graph(G)
```
